% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

\hypertarget{guide-to-descriptive-statistics-formulas-use-cases-and-assumptions}{%
\section{Guide to Descriptive Statistics: Formulas, Use Cases, and
Assumptions}\label{guide-to-descriptive-statistics-formulas-use-cases-and-assumptions}}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Descriptive statistics are fundamental tools in data analysis, providing simple
summaries about the sample and the measures. They help researchers, analysts,
and students understand the basic features of data, making complex information
more accessible and interpretable. By reducing large volumes of data into
digestible figures, descriptive statistics form the foundation for further
statistical analysis and decision-making.

\hypertarget{overview-of-descriptive-statistics}{%
\subsection{Overview of Descriptive
Statistics}\label{overview-of-descriptive-statistics}}

Descriptive statistics summarize and organize characteristics of a data set.
They are typically divided into measures of central tendency and measures of
dispersion. These statistics are essential for presenting quantitative
descriptions in a manageable form, allowing for the identification of patterns,
trends, and anomalies within the data. Descriptive statistics do not attempt to
draw conclusions beyond the data analyzed or infer causality.

\hypertarget{measures-of-central-tendency}{%
\subsection{Measures of Central Tendency}\label{measures-of-central-tendency}}

Measures of central tendency describe the center or typical value of a dataset.
The three most common measures are the mean, median, and mode.

\hypertarget{mean}{%
\subsubsection{Mean}\label{mean}}

The mean (average) is calculated as:

\[
\bar{x} = \frac{1}{n} \sum_{i=1}^{n} x_i
\]

where \(x_i\) are the data points and \(n\) is the number of observations. The
mean is sensitive to outliers and is best used with interval or ratio data. It
provides a useful summary when the data are symmetrically distributed.

\textbf{Example:} Calculating the average test score of a class to assess
overall performance.

\hypertarget{median}{%
\subsubsection{Median}\label{median}}

The median is the middle value when data are ordered from least to greatest. If
\(n\) is even, it is the average of the two middle values. The median is robust
to outliers and skewed data, making it a preferred measure when the data
distribution is not symmetrical.

\textbf{Example:} Finding the median household income in a city to represent the
typical income level, especially when a few very high incomes would skew the
mean.

\hypertarget{mode}{%
\subsubsection{Mode}\label{mode}}

The mode is the value that appears most frequently in a data set. There can be
more than one mode (bimodal or multimodal distributions) or none at all if all
values are unique. The mode is useful for categorical data where we wish to know
the most common category.

\textbf{Example:} Identifying the most common shoe size sold in a store to
inform inventory decisions.

\hypertarget{measures-of-dispersion}{%
\subsection{Measures of Dispersion}\label{measures-of-dispersion}}

Measures of dispersion describe the spread or variability within a dataset. They
help to understand how much the data deviate from the central value.

\hypertarget{range}{%
\subsubsection{Range}\label{range}}

The range is the difference between the maximum and minimum values:

\[
\text{Range} = x_{\text{max}} - x_{\text{min}}
\]

The range provides a quick sense of the spread but is highly sensitive to
outliers.

\hypertarget{variance}{%
\subsubsection{Variance}\label{variance}}

Variance measures the average squared deviation from the mean:

\[
\text{Variance} = s^2 = \frac{1}{n-1} \sum_{i=1}^{n} (x_i - \bar{x})^2
\]

Variance is useful for understanding the degree of spread in the data, but its
units are the square of the original data units.

\hypertarget{standard-deviation}{%
\subsubsection{Standard Deviation}\label{standard-deviation}}

The standard deviation is the square root of the variance:

\[
\text{Standard Deviation} = s = \sqrt{\frac{1}{n-1} \sum_{i=1}^{n} (x_i - \bar{x})^2}
\]

Standard deviation is widely used because it is expressed in the same units as
the data, making interpretation more intuitive.

\textbf{Example:} Assessing the variability of students' test scores to
determine consistency in performance.

\hypertarget{assumptions-and-limitations}{%
\subsection{Assumptions and Limitations}\label{assumptions-and-limitations}}

\begin{itemize}
\tightlist
\item
  Data should be measured at the appropriate scale (interval or ratio for mean
  and standard deviation).
\item
  Outliers can distort the mean and standard deviation, so consider using the
  median and interquartile range for skewed data.
\item
  Descriptive statistics do not infer relationships or causality; they only
  describe the data at hand.
\item
  The choice of measure depends on the data type and distribution.
\end{itemize}

\hypertarget{practical-use-cases}{%
\subsection{Practical Use Cases}\label{practical-use-cases}}

\begin{itemize}
\tightlist
\item
  Summarizing survey results (e.g., average satisfaction score) to provide a
  snapshot of respondent opinions.
\item
  Reporting demographic statistics (e.g., median age) for population studies.
\item
  Quality control in manufacturing (e.g., standard deviation of product weight)
  to monitor process consistency.
\item
  Educational assessment (e.g., mode of grades) to identify common outcomes.
\end{itemize}

\hypertarget{introduction-to-regression-analysis}{%
\subsection{Introduction to Regression
Analysis}\label{introduction-to-regression-analysis}}

Regression analysis is a statistical method used to examine the relationship
between one dependent variable and one or more independent variables. While
descriptive statistics summarize data, regression analysis allows us to model
and analyze relationships, make predictions, and infer trends.

\hypertarget{linear-regression}{%
\subsubsection{Linear Regression}\label{linear-regression}}

Linear regression models the relationship between a dependent variable \(y\) and
a single independent variable \(x\) using a straight line:

\[
y = \beta_0 + \beta_1 x + \varepsilon
\]

where \(\beta_0\) is the intercept, \(\beta_1\) is the slope, and
\(\varepsilon\) is the error term. The goal is to find the line that best fits
the data, minimizing the sum of squared residuals.

\textbf{Use Case:} Predicting a student's final exam score (\(y\)) based on the
number of hours studied (\(x\)).

\hypertarget{multiple-linear-regression}{%
\subsubsection{Multiple Linear Regression}\label{multiple-linear-regression}}

Multiple linear regression extends the concept to include two or more
independent variables:

\[
y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \cdots + \beta_p x_p + \varepsilon
\]

where \(x_1, x_2, ..., x_p\) are independent variables. This model is used when
the outcome is influenced by several factors.

\textbf{Use Case:} Predicting house prices (\(y\)) based on square footage
(\(x_1\)), number of bedrooms (\(x_2\)), and location (\(x_3\)).

\hypertarget{assumptions-of-regression-analysis}{%
\paragraph{Assumptions of Regression
Analysis}\label{assumptions-of-regression-analysis}}

\begin{itemize}
\tightlist
\item
  Linearity: The relationship between dependent and independent variables is
  linear.
\item
  Independence: Observations are independent of each other.
\item
  Homoscedasticity: The variance of residuals is constant across all levels of
  the independent variables.
\item
  Normality: Residuals are normally distributed.
\end{itemize}

Regression analysis is a powerful tool for understanding and predicting
outcomes, but it requires careful attention to its assumptions and the quality
of the data.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

Descriptive statistics provide essential tools for summarizing and understanding
data. By applying the correct measures and formulas, analysts can present clear,
concise, and meaningful insights. Regression analysis builds on these
foundations, enabling the exploration of relationships and predictions based on
data.

\hypertarget{references}{%
\subsection{References}\label{references}}

\begin{itemize}
\tightlist
\item
  Gravetter, F. J., \& Wallnau, L. B. (2017). Statistics for the Behavioral
  Sciences (10th ed.). Cengage Learning.
\item
  Urdan, T. C. (2016). Statistics in Plain English (4th ed.). Routledge.
\item
  American Psychological Association. (2020). Publication Manual of the American
  Psychological Association (7th ed.).
\item
  Field, A. (2018). Discovering Statistics Using IBM SPSS Statistics (5th ed.).
  Sage.
\item
  Montgomery, D. C., Peck, E. A., \& Vining, G. G. (2012). Introduction to
  Linear Regression Analysis (5th ed.). Wiley.
\end{itemize}

\end{document}
